---
title: "script"
author: "APB"
date: "4/30/2021"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(echo = TRUE)
```


#Load/Clean data
```{r load_data, include=FALSE}
feed_data<-read.csv("data/feeding_data_3.28.21.csv", header=T)

po<-read.csv("data/4_Day_PO.csv", header = T) #Corrected_Change is the Active PO per sample (~8 ind. per sample)

po_correct<- # sample number is arbitrary
  read_csv("data/PO_biodare_sample_correction.csv") %>%
  pivot_longer(cols = 2:5, names_sep = "_", 
               names_to = c(NA,"sample"), values_to = "PO")

growth<-read.csv("data/growth_data_3.28.21.csv", header=T)

life<-read.csv("data/life_data_3.28.21.csv", header =T)

```


#Calculate feeding rates
```{r animal_feed_rates, include = FALSE}
#calculate the mean of the controls for each plate
summary_data <- feed_data %>%
  group_by(Plate_Control) %>%
  mutate(control_mean = mean(Control_Flour_Reading, na.rm = T)) %>%
  select(Plate_Control, control_mean) %>%
  distinct() %>% as.data.frame()

#remove extra row
summary_data  <- summary_data[!(summary_data$Plate_Control == ""),]

#match control mean with plate treatment
map = setNames(summary_data$control_mean,summary_data$Plate_Control)
feed_data$control_mean <- map[as.character(feed_data$Plate_Treatment)]

#volume, mL
v = 10

#time, hours
t = 9

##difference in feeding compared to control
k=feed_data$control_mean/feed_data$Flourometry_Reading

##calculate feeding rate (Sarnelle and Wilson)
feed_data$fr_sw <- log(k) * (v/t)

#calculate mean feeding rate per animal (keep rate > 0)
animal_mean <- feed_data %>%
  #keep fr_sw values > 0
  #filter(fr_sw > 0) %>% #this removes 64 animals...
  #group based on these variables
  group_by(Time_treatment, Age, Animal, Exposed.Control) %>%
  #means based on groups above
  summarise(fr_sw = mean(fr_sw, na.rm=TRUE)) 

```


#Merge dataframes
```{r combine_dataframes, include=FALSE}
#combine feeding rates with life history data 
merge(animal_mean, life, by=c("Age", "Time_treatment", "Animal", "Exposed.Control"), keep=TRUE) %>%
  mutate(id = paste(Age,
                    Time_treatment,
                    Animal,
                    Exposed.Control, 
                    sep = "_")) %>%
           as.data.frame()->all.data


#create id for growth data 
growth %>%
  mutate(id = paste(Age,
                    Time_treatment,
                    Animal,
                    Exposed.Control, 
                    sep = "_")) %>%
           as.data.frame()->growth
# join growth data to feeding rate and life history data
all.data <- inner_join(all.data, growth, by = c("Age", "Time_treatment", "Animal", "Exposed.Control", "id"))
```

#Exploratory Plots
```{r feeding_rate_plot, echo=FALSE}
all.data %>%
  filter(Sex == "F")%>%
  ggplot()+
  geom_boxplot(aes(x=as.factor(Age), y=fr_sw, color=Time_treatment))+
  scale_color_manual(values=c("goldenrod2", "darkblue"))+
  facet_grid(.~Exposed.Control)+
  theme_bw()

all.data %>%
  filter(Sex == "F")%>%
  ggplot()+
  geom_jitter(aes(x=Time_treatment, y=fr_sw, color=Exposed.Control))+
  scale_color_manual(values=c("black", "limegreen"))+
  facet_grid(.~Age)+
  theme_bw()

```

```{r po_plots, echo=FALSE}
po %>%
  ggplot()+
  geom_point(aes(x=Hours_from_Start, y =Corrected_Change, color=Time_treatment))+
  scale_color_manual(values=c("goldenrod2", "darkblue"))+
  geom_smooth(aes(x=Hours_from_Start, y =Corrected_Change), color="black", span=1)+
  theme_bw()

po_correct %>%
  ggplot() +
  geom_point(aes(x = time, y = PO, group = sample)) +
  geom_smooth(aes(x = time, y = PO), method = "lm") +
  theme_bw()
```

```{r growth_plots, echo=FALSE}
growth %>%
  filter(Age == 6) %>%
  ggplot(aes(x=days_post_exposure, y=Length_at_exposure_mm, color=as.factor(Animal)))+
  geom_point()+
  geom_line()+
  facet_grid(Time_treatment~.)

growth %>%
  filter(Age == 7) %>%
  ggplot(aes(x=days_post_exposure, y=Length_at_exposure_mm, color=as.factor(Animal)))+
  geom_point()+
  geom_line()+
  facet_grid(Time_treatment~.)

growth %>%
  filter(Age == 8) %>%
  ggplot(aes(x=days_post_exposure, y=Length_at_exposure_mm, color=as.factor(Animal)))+
  geom_point()+
  geom_line()+
  facet_grid(Time_treatment~.)

growth %>%
  filter(Age == 9) %>%
  ggplot(aes(x=days_post_exposure, y=Length_at_exposure_mm, color=as.factor(Animal)))+
  geom_point()+
  geom_line()+
  facet_grid(Time_treatment~.)

#lag and lead

```


```{r infection_plot, echo=FALSE}

# infection prevalence
all.data %>%
  filter(Time_treatment == "Exposed" | Sex == "F") %>%
  ggplot(aes(x=Age, y=infected, color=Time_treatment))+
  scale_color_manual(values=c("goldenrod2", "darkblue"))+
  geom_jitter(width = .1, height = .1)+
  geom_smooth(method="lm")+
  scale_y_continuous(limits=c(-.1,1))+
  scale_x_continuous(breaks=c(6,7,8,9))+
  theme_bw()


#infection intensity
life %>%
  filter(Time_treatment == "Exposed"| Sex == "F") %>%
  ggplot(aes(x=Age, y=spore_counts, color=Time_treatment))+
  scale_color_manual(values=c("goldenrod2", "darkblue"))+
  #geom_jitter(width=.25)+
  geom_point()+
  geom_smooth(method="lm")+
  theme_bw()

```


```{r offsrping_plots, echo=FALSE}

#total offspring
all.data %>%
  filter(Sex == "F") %>%
  #filter(infected != "NA") %>%     #removes controls
  ggplot(aes(x=Time_treatment, y=total_offspring, fill=as.factor(infected)))+
  geom_boxplot()+
  facet_grid(~Age)


#age at first clutch
all.data %>% 
  filter(Sex == "F") %>%
  #filter(infected != "NA") %>%     #removes controls
  ggplot(aes(x=Time_treatment, y=age_at_first_clutch, fill=as.factor(infected)))+
  geom_boxplot()+
  facet_grid(~Age)

```


#Stats

# PO
```{r po}

#Does PO differ across a day (Resting = Light hours, Active = dark hours) and across ages? 
summary(po_model<-glm(Corrected_Change ~ Phase*as.factor(Age), data=po))
plot(po_model)

summary(lm(Corrected_Change ~ as.factor(Age), data = po))
plot(lm(Corrected_Change ~ Age, data = po))

newdata<-expand.grid(Phase=c("Active", "Resting"), Age = 6:9)
newdata$preds<-predict(po_model, newdata=newdata)

newdata %>%
  ggplot()+
  geom_line(aes(x=Age, y=preds, color=Phase))+
  geom_point(po, mapping=aes(x=Age, y=Corrected_Change, color=Phase), alpha = 0.5)
```

```{r}
hist(po$Corrected_Change)

# po is strictly? positive
# no zeros
po_summary <- po %>%
  group_by(Age) %>%
  mutate(age_mean_po = mean(Corrected_Change, na.rm = TRUE)) %>%
  group_by(Time_treatment, Age) %>%
  summarise(mean_po = mean(Corrected_Change, na.rm = TRUE),
            min_po = min(Corrected_Change, na.rm = TRUE),
            max_po = max(Corrected_Change, na.rm = TRUE),
            age_mean_po = unique(age_mean_po))
po_summary

```

```{r}
# linear model using log(Corrected_Change) as response
fit.log.po <- lm(log(Corrected_Change) ~ Age * Time_treatment, data = po)
summary(fit.log.po)
plot(fit.log.po)

# plot predictions
nd <- expand.grid(Time_treatment=c("Dark", "Light"), Age = seq(6, 9, length.out = 100))
po_pred <- predict(fit.log.po, newdata = nd, interval = "confidence") %>%
  as_tibble() %>%
  mutate(fit_median = exp(fit),
         lwr_median = exp(lwr),
         upr_median = exp(upr)) %>%
  cbind(nd)

po_pred %>%
ggplot() +
  geom_line(aes(x = Age, y = fit_median, col = Time_treatment)) +
  geom_ribbon(aes(x = Age, ymin = lwr_median, ymax = upr_median, fill = Time_treatment), alpha = 0.3) +
  geom_point(data = po_summary, mapping = aes(x = Age, y = mean_po, col = Time_treatment)) +
  geom_point(data = po_summary, mapping = aes(x = Age, y = age_mean_po))
  #geom_jitter(data = po, mapping = aes(x = Age, y = Corrected_Change, col = Time_treatment), width = 0.1, height = 0)

glm(Corrected_Change ~ Age * Time_treatment, family = gaussian(link = "log"), data = po)
```


# infection



```{r infection_prevalence}
prev <- all.data %>%
  filter(Exposed.Control == "Exposed", 
         days_post_exposure == 0) %>%
  group_by(Age, Time_treatment) %>%
  summarise(inf = sum(infected, na.rm=T),
            n = n(),
            prev = inf/n)

prev %>% knitr::kable()

prev %>% ggplot() +
  geom_point(aes(x = Age, y = prev, col = Time_treatment)) +
  geom_line(aes(x = Age, y = prev, col = Time_treatment)) +
  ylim(0,1)
```

```{r}
# get subset for infection and spore count analysis
# arbitrarily choose 0 days post exposure since the data is duplicated for each days_post_exposure
reg_data <- filter(all.data, Exposed.Control == "Exposed", days_post_exposure == 0) 

# logistic regression
# model 1: Time_treatment only
inf_time <- glm(infected ~ Time_treatment, family = binomial , data = reg_data)
# model 2: common linear effect of age with Time_treatment specific intercepts
inf_main <- glm(infected ~ Age + Time_treatment, family = binomial, data = reg_data)
# model 3: Time_treatment specific intercept and slopes
inf_full_lin <- glm(infected ~ Age * Time_treatment, family = binomial, data = reg_data)
# model 4: full interaction model with age as categorical
inf_full_cat <- glm(infected ~ factor(Age) * Time_treatment, family = binomial , data = reg_data)

# 
summary(inf_time)
summary(inf_main)
summary(inf_full_lin)
summary(inf_full_cat)

# residual plots
plot(inf_time, 1)
plot(inf_main, 1)
plot(inf_full_lin, 1)
plot(inf_full_cat, 1)
```

Based on the sign of the coefficient for Time_treatment, all models agree that night-time exposure increases the probability of infection. The model with just time of exposure is pretty good compared to the others. Based on the likelihood ratio tests and AIC comparison below, I think that if we were going to choose one model that we should go with model 2 (infected ~ Age + time_treatment).

Note: We should try including age at death as a predictor of infection (if it makes sense, i.e., dying soon after exposure "censors" the infection outcome) or spore counts. Nevermind, these data are already coded as NA if daphnia die before age 8.

```{r}
# likelihood ratio test for significance of linear age term
# marginally significant
car::Anova(inf_main, test = "LR")

# likelihood ratio test for significance of exposure time specific effect of age
# Not significant
car::Anova(inf_full_lin, test = "LR")

```

```{r}
AIC(inf_time, inf_main, inf_full_lin, inf_full_cat) %>%
  mutate(deltaAIC = AIC - min(AIC)) %>% add_column(model = c("time", "Age + time", "Age*time", "factor(Age)*time"), .before = 1)
```


```{r}
crit <- 1.96

inf_preds <- 
  predict(inf_main, newdata = prev, type = "link", se.fit = TRUE) %>% 
  as_tibble() %>%
  mutate(half = crit * se.fit,
         lwr = fit - half,
         upr = fit + half,
         fit = boot::inv.logit(fit),
         lwr = boot::inv.logit(lwr),
         upr = boot::inv.logit(upr)) %>%
  cbind(prev)

inf_preds %>%
  ggplot() +
  geom_line(aes(x = Age, y = fit, col = Time_treatment, linetype = Time_treatment)) +
  geom_ribbon(aes(x = Age, ymin = lwr, ymax = upr, fill = Time_treatment), alpha = 0.3) +
  geom_point(aes(x  = Age, y = prev, col = Time_treatment, shape = Time_treatment), size = 2) +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  ylim(0,1) +
  labs(y = "infection probability",
       x = "Age at exposure (days)",
       col = "Exposure\ntime",
       shape = "Exposure\ntime",
       linetype = "Exposure\ntime",
       fill = "Exposure\ntime") +
  theme_bw()
```

Comparison of model predicted probability of infection plus 95% Wald CI and the observed prevalence. It is clear that the probability of infection is greater during the night than the day. There is some evidence that the probability of infection increases with the age of exposure.

```{r}
# make table of estimated probabilities and 95% CI
inf_preds %>% select(Age, Time_treatment, inf, n, prev, fit, lwr, upr) %>%
  group_by(Age, Time_treatment) %>%
  transmute("observed prevalence" = paste0(round(prev, 2), " (", inf, "/" , n, ")"),
         "estimated prevelence" = paste0(round(prev,2), " (", round(lwr,2), ", ", round(upr,2), ")")) %>%
  knitr::kable()
  
```


```{r eval = FALSE}
library(mcprofile)
K <- matrix(c(1, 6, 0,
              1, 6, 1,
              1, 7, 0,
              1, 7, 1,
              1, 8, 0,
              1, 8, 1,
              1, 9, 0,
              1, 9, 1), byrow = TRUE, nrow = 8, ncol = 3)

lin.combo <- mcprofile(inf_main, CM = K)
# get CI for log odds
inf_CI_unadjusted <- confint(lin.combo, level = 0.95, adjust = "none")
inf_CI_bonferroni <- confint(lin.combo, level = 0.95, adjust = "bonferroni")

# transform estimate and interval to probability scale
inf_CI_lr_unadjusted <-  cbind(inf_CI_unadjusted$estimate, inf_CI_unadjusted$confint) %>%
  mutate(prob = boot::inv.logit(Estimate),
         prob_lwr = boot::inv.logit(lower),
         prob_upr = boot::inv.logit(upper),
         adjust = "none") %>%
  cbind(prev)

inf_CI_lr_bonferroni <- cbind(inf_CI_bonferroni$estimate, inf_CI_bonferroni$confint) %>%
  mutate(prob = boot::inv.logit(Estimate),
         prob_lwr = boot::inv.logit(lower),
         prob_upr = boot::inv.logit(upper),
         adjust = "Bonferroni") %>%
  cbind(prev)

inf_lr_CI <- rbind(inf_CI_lr_bonferroni, inf_CI_lr_unadjusted)
inf_lr_CI %>% knitr::kable(digits = 2)
```

We are 95% confident that the odds of infection when Daphnia are exposed at night is 1.31 to 2.76 times as large when exposed at day when age of exposure is held constant. We are 95% confident that the odds of infection 

```{r eval = FALSE}
inf_lr_CI %>%
  ggplot() +
  geom_line(aes(x = Age, y = prob, col = Time_treatment)) +
  geom_ribbon(aes(x = Age, ymin = prob_lwr, ymax = prob_upr, fill = Time_treatment), alpha = 0.3) +
  geom_point(aes(x = Age, y = prev, col = Time_treatment)) +
  facet_wrap(~adjust)
```
```{r eval = FALSE}
K <- matrix(c(0, 1, 0,
              0, 0, 1), 
            byrow = TRUE, nrow = 2, ncol = 3)

lin.combo <- mcprofile(inf_main, CM = K)
# get CI for log odds
inf_OR_unadjusted <- confint(lin.combo, level = 0.95, adjust = "none")
inf_OR_bonferroni <- confint(lin.combo, level = 0.95, adjust = "bonferroni")

inf_OR_unadjusted

```

```{r}
# built-in way to get LR confidence intervals
CI_OR_night <- confint(inf_main, parm = "Time_treatmentNight", level = 0.95)
CI_OR_day <- confint(inf_main, parm = "(Intercept)", level = 0.95)
CI_OR_age <- confint(inf_main, parm = "Age", level = 0.95)

# odds of infection by 1 day increase in age at exposure
exp(coef(inf_main)[2])
exp(CI_OR_age)

# odds of infection at night vs day
exp(coef(inf_main)[3])
exp(CI_OR_night)

# transform to probability scale
# point estimate and  95 % likelihood ratio interval
# day
boot::inv.logit(coef(inf_main)[1]) 
boot::inv.logit(CI_OR_day)
# night
boot::inv.logit(coef(inf_main)[3])
boot::inv.logit(CI_OR_night)


```

What if we looked at the random effect of age instead of fixed linear effect? It may be the case that the increase is infection probability is driven by the age 6 day group which has only one infected daphnia. There may be a random effect of age since we see a similar up-down pattern across ages for each time treatment.

```{r}
library(lme4)

# fit binomial model with random intercept by age grouping
# bernoulli model
inf_rand_age <- glmer(infected ~ Time_treatment + (1 | Age), family = binomial(), data = reg_data)
# binomial
prev_rand_age <- glmer(inf/n ~ Time_treatment + (1 | Age), family = binomial(), data = prev, weights = n)
# fit binomial glm with to compare with binomial mixed model
prev_main <- glm(inf/n ~ Time_treatment + Age, family = binomial(), data = prev, weights = n)
# random effect estimates are similar between bernoulli and binomial specification
summary(inf_rand_age)
summary(prev_rand_age)

# compare AIC
AIC(inf_main, inf_rand_age)
AIC(prev_main, prev_rand_age)
```


# spore intensity

```{r}
zero_summary <- 
  reg_data %>%
  group_by(Age, Time_treatment) %>%
  mutate(conditional_spore = ifelse(infected == 0, NA, spore_counts)) %>%
  summarise(mean_spore_counts = mean(spore_counts, na.rm = TRUE),
            var_spore_counts = var(spore_counts, na.rm = TRUE),
            cond_mean_spore_counts = mean(conditional_spore, na.rm = TRUE),
            n_zeros = sum(spore_counts == 0, na.rm = TRUE),
            n_na = sum(is.na(spore_counts)),
            n_infected = sum(infected == 1, na.rm = TRUE),
            n_animals = n())
zero_summary %>% knitr::kable(digits = 2)

# 5 daphnia have just 1 spore counted, including the only daphnia that was infected in Age 6 Day trt.
reg_data %>%
  filter(spore_counts == 1)
  
```

There are a good deal of zeros. The variance is much bigger than the mean in many treatment combinations. Zero-inflated negative binomial makes sense here.

```{r}
ggplot() +
  geom_jitter(data = filter(reg_data, Time_treatment == "Day"), 
              aes(x = Age - 0.1, y = spore_counts, col = Time_treatment), height = 0, width = 0.1) +
  geom_jitter(data = filter(reg_data, Time_treatment == "Night"), 
              aes(x = Age + 0.1, y = spore_counts, col = Time_treatment), height = 0, width = 0.1) +
  geom_point(data = filter(zero_summary, Time_treatment == "Day"), 
             aes(x = Age - 0.1, y = mean_spore_counts, col = Time_treatment), 
             size = 4, shape = 3) +
  geom_point(data = filter(zero_summary, Time_treatment == "Night"), 
             aes(x = Age + 0.1, y = mean_spore_counts, col = Time_treatment), 
             size = 4, shape = 3) +
      geom_point(data = filter(zero_summary, Time_treatment == "Day"), 
             aes(x = Age - 0.1, y = cond_mean_spore_counts, col = Time_treatment), 
             size = 4, shape = 2) +
  geom_point(data = filter(zero_summary, Time_treatment == "Night"), 
             aes(x = Age + 0.1, y = cond_mean_spore_counts, col = Time_treatment), 
             size = 4, shape = 2) 
```

There appears to be an age effect in spore counts. Individuals that were infected at age 9 have a greater mean spore count than individuals infected at day 6 (all Daphnia were followed up for 12 days post-exposure, i.e., for age 6 individuals their spores were counted at age 18 and age 9 daphnia were counted at age 21). This might be caused by size differences between different aged individuals since maybe daphnia that are bigger over the course of the 12 days that spores could develop would provide more energy for spore production. We could try to account for this using body size (length or length^3^ (since feeding seems to correlate with surface area)) at time of exposure as a baseline (The reg_data set already has this measurement).

Sidebar: it may be interesting to try and include the growth curves as a functional covariate since maybe it is a proxy for how much total energy is available to spore production over the interval we allow them to grow.

```{r}
# Per pscl vignette suggestions, lets look at the log count 
# since we use log link for count model
# log(x + 0.5)
clog <- function(x) log(x + 0.5)

ggplot() +
  geom_jitter(data = filter(reg_data, Time_treatment == "Day"), 
              aes(x = Age - 0.1, y = clog(spore_counts), col = Time_treatment), height = 0, width = 0.1) +
  geom_jitter(data = filter(reg_data, Time_treatment == "Night"), 
              aes(x = Age + 0.1, y = clog(spore_counts), col = Time_treatment), height = 0, width = 0.1)  +
    geom_point(data = filter(zero_summary, Time_treatment == "Day"), 
             aes(x = Age - 0.1, y = clog(mean_spore_counts), col = Time_treatment), 
             size = 4, shape = 22) +
  geom_point(data = filter(zero_summary, Time_treatment == "Night"), 
             aes(x = Age + 0.1, y = clog(mean_spore_counts), col = Time_treatment), 
             size = 4, shape = 22) +
      geom_point(data = filter(zero_summary, Time_treatment == "Day"), 
             aes(x = Age - 0.1, y = clog(cond_mean_spore_counts), col = Time_treatment), 
             size = 4, shape = 24) +
  geom_point(data = filter(zero_summary, Time_treatment == "Night"), 
             aes(x = Age + 0.1, y = clog(cond_mean_spore_counts), col = Time_treatment), 
             size = 4, shape = 24) 


#plot(clog(spore_counts) ~ factor(interaction(Time_treatment, factor(Age))), data = reg_data)
```

Looking at the log count, it actually appears that only the day treatment has an positive effect of age on spore count. There doesn't appear to be any age effect on mean spore count for the night treatment.

```{r}
# look for potential colinearity
reg_data %>%
  ggplot() +
  geom_jitter(aes(x = Age, y = Length_at_exposure_mm, col = factor(infected)))

reg_data %>%
  ggplot() +
  geom_jitter(aes(x = Age, y = Length_at_exposure_mm^3, col = factor(infected)))

cor.test(reg_data$Age, reg_data$Length_at_exposure_mm)
```


```{r}
reg_data %>%
  ggplot() +
  geom_point(aes(x = Length_at_exposure_mm, y = clog(spore_counts)))

reg_data %>%
  ggplot() +
  geom_point(aes(x = Length_at_exposure_mm^3, y = clog(spore_counts)))

```

Hmm, maybe length at exposure isn't that informative? Ignoring the zero counts, it appear that log(spore counts) doesn't vary with length of volume of daphnia at exposure.

### zero-inflated and hurdle models

It's not clear to me how to evaluate the fit of the these models. Need to figure out. DHARMa?

```{r}
#library(pscl)
library(countreg)
library(lmtest)

```

```{r eval = FALSE}
# specify a mess of models
model_formulas <- 
  c(spore_counts ~ 1 | 1,
    spore_counts ~ 1 | Age + Time_treatment,
    spore_counts ~ 1 | Time_treatment,
    spore_counts ~ Time_treatment | Time_treatment,
    spore_counts ~ Time_treatment | Age + Time_treatment, 
    spore_counts ~ Time_treatment*Age | Age + Time_treatment,
    spore_counts ~ Time_treatment*Age | Time_treatment,
                  spore_counts ~ Time_treatment + Age | Age + Time_treatment, 
                  spore_counts ~ Time_treatment * Age | Age + Time_treatment, 
                  spore_counts ~ Time_treatment + Length_at_exposure_mm | Age + Time_treatment, 
                  spore_counts ~ Time_treatment + poly(Length_at_exposure_mm, 3) | Age + Time_treatment,
                  spore_counts ~ Time_treatment + poly(Length_at_exposure_mm, 3) + Age | Age + Time_treatment,
                  spore_counts ~ Time_treatment + Length_at_exposure_mm + Age | Age + Time_treatment)

# fit all models
mod_list <- lapply(seq_along(model_formulas), function(x) hurdle(model_formulas[[x]],
                                                                 dist = "negbin", link = "logit", data = reg_data))

# print AIC
map_dbl(mod_list, AIC)

lrtest(mod_list[[2]], mod_list[[3]]) # test significant of Age in zero hurdle part
lrtest(mod_list[[3]], mod_list[[4]]) # test significance of Time_treatment in count part
```


```{r}
# ZI models 
# negative binomial
zi_nb_null <- zeroinfl(spore_counts ~ 1 | Age + Time_treatment,
       dist = "negbin", link = "logit", data = reg_data)
zi_nb_time <- zeroinfl(spore_counts ~ Time_treatment | Age + Time_treatment,
       dist = "negbin", link = "logit", data = reg_data)
zi_nb_age_time <- zeroinfl(spore_counts ~ Age + Time_treatment | Age + Time_treatment,
       dist = "negbin", link = "logit", data = reg_data)
zi_nb_age_time_int <- zeroinfl(spore_counts ~ Age * Time_treatment | Age + Time_treatment,
       dist = "negbin", link = "logit", data = reg_data)
# poisson
zi_poi_null <- zeroinfl(spore_counts ~ 1 | Age + Time_treatment,
       dist = "poisson", link = "logit", data = reg_data)
zi_poi_time <- zeroinfl(spore_counts ~ Time_treatment | Age + Time_treatment,
       dist = "poisson", link = "logit", data = reg_data)
zi_poi_age_time <- zeroinfl(spore_counts ~ Age + Time_treatment | Age + Time_treatment,
       dist = "poisson", link = "logit", data = reg_data)
zi_poi_age_time_int <- zeroinfl(spore_counts ~ Age * Time_treatment | Age + Time_treatment,
       dist = "poisson", link = "logit", data = reg_data)

# Hurdle models
# # negative binomial
# hurd_nb_null <- hurdle(spore_counts ~ 1 | Age + Time_treatment,
#        dist = "negbin", link = "logit", data = reg_data)
# hurd_nb_time <- hurdle(spore_counts ~ Time_treatment | Age + Time_treatment,
#        dist = "negbin", link = "logit", data = reg_data)
# hurd_nb_age_time <- hurdle(spore_counts ~ Age + Time_treatment | Age + Time_treatment,
#        dist = "negbin", link = "logit", data = reg_data)
# hurd_nb_age_time_int <- hurdle(spore_counts ~ Age * Time_treatment | Age + Time_treatment,
#        dist = "negbin", link = "logit", data = reg_data)
# # poisson
# hurd_poi_null <- hurdle(spore_counts ~ 1 | Age + Time_treatment,
#        dist = "poisson", link = "logit", data = reg_data)
# hurd_poi_time <- hurdle(spore_counts ~ Time_treatment | Age + Time_treatment,
#        dist = "poisson", link = "logit", data = reg_data)
# hurd_poi_age_time <- hurdle(spore_counts ~ Age + Time_treatment | Age + Time_treatment,
#        dist = "poisson", link = "logit", data = reg_data)
# hurd_poi_age_time_int <- hurdle(spore_counts ~ Age * Time_treatment | Age + Time_treatment,
#        dist = "poisson", link = "logit", data = reg_data)

bbmle::AICtab(zi_nb_null,
              zi_nb_time,
              zi_nb_age_time,
              zi_nb_age_time_int,
              zi_poi_null,
              zi_poi_time,
              zi_poi_age_time,
              zi_poi_age_time_int,
              # hurd_nb_null,
              # hurd_nb_time,
              # hurd_nb_age_time,
              # hurd_nb_age_time_int,
              # hurd_poi_null,
              # hurd_poi_time,
              # hurd_poi_age_time,
              # hurd_poi_age_time_int
              logLik = TRUE, base = TRUE) %>%
  data.frame() %>%
  dplyr::select(-dLogLik) %>%
  # write.csv("aic_spore_count.csv")
  knitr::kable(digits = 2)
```

So, we can see from AIC comparison that the negative binomial is better than poisson.

Let's look at likelihood ratio tests for model selection.
```{r}
# using function from lmtest
lrtest(zi_nb_null, zi_nb_time)
lrtest(zi_nb_time, zi_nb_age_time)
lrtest(zi_nb_age_time, zi_nb_age_time_int)
```


```{r}
# is this correct?
# output agrees with previous code chunk
car::Anova(zi_nb_age_time_int, test = "Chisq")
```

Both the pair-wise lrtest and the car::Anova test provide the same qualitative conclusions. None of the terms are statistically significant, i.e., the the probability that we would observe the improvement in model fit (likelihood) that is obtained by adding explanatory variables when those explanatory variables don't matter is large. There do not appear to be any differences in the expected count among treatments after accounting for differences in infection probability.

```{r}
library(emmeans)

predict(zi_nb_time, newdata = expand.grid(Time_treatment = c("Day", "Night"), Age = 6:9), type = "response")
predict(zi_nb_time, newdata = expand.grid(Time_treatment = c("Day", "Night"), Age = 6:9), type = "count")
predict(zi_nb_time, newdata = expand.grid(Time_treatment = c("Day", "Night"), Age = 6:9), type = "zero")

# zi_nb_time
emm_response <- emmeans(zi_nb_time, "Time_treatment", mode = "response") # arithmatic mean
emm_response_lin <- emmeans(zi_nb_time, "Time_treatment", mode = "response", lin.pred = T) # geometric mean
emm_count <- emmeans(zi_nb_time, "Time_treatment", mode = "count")
# emmeans(zi_nb_time, "Time_treatment", mode = "prob0")

# zi_nb_age_time
emm_response_age_time <- emmeans(zi_nb_age_time, specs = "Time_treatment", by = "Age", mode = "response") # arithmatic mean
emm_count_age_time <- emmeans(zi_nb_age_time, specs = "Time_treatment", by = "Age", mode = "count")
```

```{r}
# get CI for plotting response
model_mat <- model.matrix(zi_nb_age_time) %>% unique()
rownames(model_mat) <- c("age6_day", "age6_night",
                      "age7_day", "age7_night",
                      "age8_day", "age8_night",
                      "age9_day", "age9_night")
model_mat

contr_list <- lapply(1:nrow(model_mat), function(x) model_mat[x,])
names(contr_list) <- rownames(model_mat)

contrast(emm_response_age_time, spec = "pairwise", by = "Age")

emmip(emm_response_age_time, ~ Age | Time_treatment, CIs = TRUE) +
    geom_point(aes(x = Age, y = spore_counts), data = reg_data, pch = 2, color = "blue")
```


```{r}
# mod.mat <- model.matrix(zi_nb_time) %>% unique()
# row.names(mod.mat) <- c("day", "night")
# 
# night_vs_day <- list("night vs day, age 6" = mod.mat["night",] - mod.mat["day",])

# response
# lf_response <- contrast(emm_response, night_vs_day)
# cont_response <- contrast(emm_response, "revpairwise")
# summary(lf_response)
# confint(lf_response)
# 

summary(contrast(emm_response, "revpairwise"))
confint(contrast(emm_response, "revpairwise")) # night - day for response
# confint(contrast(emm_response_lin, "revpairwise")) # using linear.predictor scale doesn't change anything

# count
cont_count <- contrast(emm_count, "revpairwise") # night - day for count
summary(cont_count)
confint(cont_count)
```

```{r}
nd_time <- zero_summary %>% dplyr::select(Age, Time_treatment)
emm_count_df <- emm_count %>% data.frame() %>% rename("count_time" = emmean, "count_lcl" = asymp.LCL, "count_ucl" = asymp.UCL)
emm_response_df <- emm_response %>% data.frame() %>% rename("response_time_emm" = emmean, "response_lcl" = asymp.LCL, "response_ucl" = asymp.UCL)

nd_time$count_time <- predict(zi_nb_time, newdata = nd_time, type = "count")
nd_time$response_time <- predict(zi_nb_time, newdata = nd_time, type = "response")

nd_time <- inner_join(nd_time, emm_count_df, by = c("Time_treatment", "count_time")) 
nd_time <- inner_join(nd_time, emm_response_df, by = c("Time_treatment"))

# plot predicted and observed counts
ggplot() +
  geom_ribbon(data = nd_time, 
              aes(x = Age, ymin = count_lcl, ymax = count_ucl, fill = Time_treatment), 
              alpha = 0.4) +  
  geom_jitter(data = zero_summary, 
             aes(x = Age, y = cond_mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 1, height = 0, width = 0.1) +
  geom_jitter(data = filter(reg_data, infected == 1), 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
  geom_line(data = nd_time, 
            aes(x = Age, y = count_time, col = Time_treatment, linetype = Time_treatment)) +
    labs(y = "Spore counts when infected", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       fill = "Exposure\ntime",
       linetype = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw()

# plot predicted and observed response
# emm output with CI
ggplot() +
  geom_ribbon(data = nd_time, 
              aes(x = Age, ymin = response_lcl, ymax = response_ucl, fill = Time_treatment), 
              alpha = 0.4) +  
  geom_jitter(data = zero_summary, 
             aes(x = Age, y = mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 1, height = 0, width = 0.1) +
  geom_jitter(data = reg_data, 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
  geom_line(data = nd_time, 
            aes(x = Age, y = response_time_emm, col = Time_treatment, linetype = Time_treatment)) +
    labs(y = "Spore counts", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       fill = "Exposure\ntime",
       linetype = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw()

# predict output no CI
ggplot() +
  geom_jitter(data = zero_summary, 
             aes(x = Age, y = mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 1, height = 0, width = 0.1) +
  geom_jitter(data = reg_data, 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
  geom_line(data = nd_time, 
            aes(x = Age, y = response_time, col = Time_treatment, linetype = Time_treatment)) +
  geom_line(data = nd_time, 
            aes(x = Age, y = response_time_emm, linetype = Time_treatment)) +
    labs(y = "Spore counts", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       fill = "Exposure\ntime",
       linetype = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw()

```

In the last plot, we can clearly see that the emmeans package computed the count based on the average age at exposure since the lines cross at 7.5.

```{r}
# here is how to specify level of continuous predictor
# spore_counts ~ time
emm_response_time <- emmeans(zi_nb_age_time, ~ Time_treatment | Age, at = list(Age = 6:9))
(emm_response_time_df <- emm_response_time %>% data.frame())
# spore_counts ~ age + time
emm_response_age_time <- emmeans(zi_nb_time, ~ Time_treatment | Age, at = list(Age = 6:9))
(emm_response_age_time_df <- emm_response_age_time %>% data.frame())

# contrast between day and night
contrast(emm_response_time, "pairwise")
```

```{r}
p_count <- 
  ggplot() +
  geom_ribbon(data = nd_time, 
              aes(x = Age, ymin = count_lcl, ymax = count_ucl, fill = Time_treatment), 
              alpha = 0.3) +  
  geom_jitter(data = zero_summary, 
             aes(x = Age, y = cond_mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 1, height = 0, width = 0.1) +
  geom_jitter(data = filter(reg_data, infected == 1), 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
  geom_line(data = nd_time, 
            aes(x = Age, y = count_time, col = Time_treatment, linetype = Time_treatment)) +
    labs(y = "Spore counts when infected", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       fill = "Exposure\ntime",
       linetype = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw()


p_response <- 
  ggplot() +
  geom_line(data = emm_response_time_df, aes(x = Age, y = emmean, col = Time_treatment, linetype = Time_treatment)) +
  geom_ribbon(data = emm_response_time_df, 
              aes(x = Age, ymin = asymp.LCL, ymax = asymp.UCL, fill = Time_treatment), alpha = 0.3) + 
  geom_jitter(data = zero_summary, 
             aes(x = Age, y = mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 1, height = 0, width = 0.1) +
  geom_jitter(data = reg_data, 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
    labs(y = "Spore counts", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       fill = "Exposure\ntime",
       linetype = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw()
  
    
```


```{r}
combined <- (p_count & theme(legend.position = "none")) + p_response
combined + plot_layout(guides = "collect") + plot_annotation(tag_levels = 'A')
```


```{r}
# zi negbin null model conditional count predictions
nd <- zero_summary %>% dplyr::select(Age, Time_treatment)
#summary(zi_nb_null)
nd$count_null <- predict(zi_nb_null, newdata = nd, type = "count")
zi_nb_null_CI <- confint(zi_nb_null) %>% as_tibble()

nd$count_null_lwr <- zi_nb_null_CI[1,1] %>% pull() %>% exp()
nd$count_null_upr <- zi_nb_null_CI[1,2] %>% pull() %>% exp()

# plot the conditional observed count, mean count, and predicted count + CI
ggplot() + 
  geom_point(data = zero_summary, 
             aes(x = Age, y = cond_mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 0.5) +
  geom_jitter(data = filter(reg_data, infected == 1), 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
  geom_line(data = nd, 
            aes(x = Age, y = count_null)) +
  geom_ribbon(data = nd, 
              aes(x = Age, ymin = count_null_lwr, ymax = count_null_upr), 
              alpha = 0.2) +
    labs(y = "Spore counts when infected", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw()

# zi negbin null model response prediction
nd$response_null <- predict(zi_nb_null, newdata = nd, type = "response")

# need to figure out how to compute CI for this

# plot mean spore count conditional on infection probability
ggplot() + 
  geom_point(data = zero_summary, 
             aes(x = Age, y = mean_spore_counts, col = Time_treatment, size = n_infected), 
             alpha = 0.5) +
  geom_jitter(data = reg_data, 
              aes(x = Age, y = spore_counts, col = Time_treatment), 
              shape = 3, height = 0, width = 0.1) +
  geom_line(data = nd, 
            aes(x = Age, y = response_null, col = Time_treatment)) +
  labs(y = "Spore counts", x = "Age at exposure (days)",
       col = "Exposure\ntime",
       size = "# infected") +
  scale_color_manual(values=c("#ffcc33", "#0033cc"))+
  scale_fill_manual(values=c("#ffcc33", "#0033cc"))+
  theme_bw() 


```


```{r}
zero_summary$response <- predict(zi_nb_time, newdata = zero_summary, type = "response")
zero_summary$count <- predict(zi_nb_time, newdata = zero_summary, type = "count")
zero_summary$prob_zero <- predict(zi_nb_time, newdata = zero_summary, type = "zero")

zero_summary %>% knitr::kable(digits = 2)
```

```{r}
zero_summary %>%
  group_by(Age) %>%
  summarise("night - day" = response - lag(response)) %>%
  ungroup() %>%
  summarise("night - day" = mean(`night - day`, na.rm = T))
``` 

